
---
title: "HW5"
author: "Junyu Wang"
output: word_document
---

## First inspection of the data

```{r}
source("~/personal/school/stats133/stat133/HW5/data_clean_preprocess.R")

str(raw_weather_data)
head(raw_weather_data)
```

## Weather Information

Remove % in column humidity and convert such values to numeric format

```{r}
head(raw_weather_data$humidity)
```

Extract the temperature from column weather, and create a column temperature2 with these values.

```{r}
head(raw_weather_data$temperature2)
```

Extract the humidity values from column weather, and create a column humidity2 with these values. 

```{r}
head(raw_weather_data$humidity2)
```

Extract the wind speed values from column weather, and create a column wind2 with these values.

```{r}
head(raw_weather_data$wind2)
```

Check new columns coincide with the pre-existing ones:

```{r}
summary(raw_weather_data$temperature)
summary(raw_weather_data$temperature2)
summary(raw_weather_data$humidity)
summary(raw_weather_data$humidity2)
summary(raw_weather_data$wind_mph)
summary(raw_weather_data$wind2)
```

## Data information

Create a column year that contains the number of the year.

```{r}
head(raw_weather_data$year)
```

Create a column monthnum that contains the number of the month.

```{r}
head(raw_weather_data$monthnum)
```

Create a column month that contains the name of the corresponding month as factor.

```{r}
head(raw_weather_data$month)
```

Create a column decade that indicates the corresponding decade of each played game.

```{r}
head(raw_weather_data$decade)
``` 

## Scores Information

Create a column total_score that contains the total number of scored points in each game.

```{r}
head(raw_weather_data$total_score)
``` 

Create a column diff_score that indicates the difference of home_score and away_score.

```{r}
head(raw_weather_data$diff_score)
``` 

Create a column home_win that shows whether home_score is greater than away_score.

```{r}
head(raw_weather_data$home_win)
``` 

## Basic Exploration

Inspect variables home_score, away_score, temperature, wind_mpg

```{r}
summary(raw_weather_data$home_score)
summary(raw_weather_data$away_score)
summary(raw_weather_data$temperature)
summary(raw_weather_data$wind_mph)
``` 

Visually inspect variables home_score, away_score, temperature, humidity and wind_mph.

```{r}
hist(raw_weather_data$home_score)
hist(raw_weather_data$away_score)
hist(raw_weather_data$temperature)
hist(raw_weather_data$humidity)
hist(raw_weather_data$wind_mph)
``` 

What team has the maximum home score?

```{r}
team_with_max_home_score
``` 

What team has the maximus away score?

```{r}
team_with_max_away_score
``` 

What is the most common home score?

```{r}
m_common_home_score
``` 

What is the most common away score?

```{r}
m_common_away_score
``` 

What has been the maximum temperature in a game?

```{r}
max_temp
``` 

What was the date of the maximum temperature?

```{r}
date_max_temp
``` 

What has been the minimum temperature in a game?

```{r}
min_temp
``` 

What was the date of the minimum temperature?

```{r}
date_min_temp
``` 

How many games have been played with a temperature of 90 degrees or more?

```{r}
games_with_high_temp
``` 

How many games have been played with a temperature below 0 degrees (do not include 0)?

```{r}
games_with_low_temp
``` 

What is the most common temperature?

```{r}
m_common_temp
``` 

Make a bar chart with the frequency table of temperatures. Is there anything that catches your attention?

```{r}
barplot(temp_freq)
``` 

The frequency of temperature 72 is much higher than the rest of temperatures. 

## Data Files

Export data for all games with selected columns.

```{r}
head(cleaned_data)
``` 

Export data for games during each decade with selected columns.

```{r}
print("files are correctly generated")
``` 

## Data Analysis

```{r}
source("~/personal/school/stats133/stat133/HW5/data_analysis.R")
``` 

## 1. Does play at home really have an advantage for the home team?

## Number of More Home victories per year

```{r}
ggplot(data = result, aes(years, diffs)) +
  geom_bar(color = "#000000", fill = "#515252", stat = "identity") +
  ggtitle("Difference between home wins and home loses by year") +
  ylab("Difference") +
  scale_x_continuous(name = "Year",
                     breaks = years) +
  theme(axis.text.x = element_text(angle= 90))
``` 

## Average score points per year

```{r}
ggplot(data = avgs) + 
  geom_line(aes(x = years, y = avg_homescore), color = "#DB5FE4", size = 1.5) +
  geom_line(aes(x = years, y = avg_awayscore), color = "#5FE4DB", size = 1.5) +
  geom_point(aes(x = years, y = avg_homescore), color = "#DB5FE4", size = 3.0) + 
  geom_point(aes(x = years, y = avg_awayscore), color = "#5FE4DB", size = 3.0) + 
  ggtitle("Average score points per year") + 
  ylab("scores")
```

Other type of analysis for this conclusion: 
We can also conduct a t-test for null hypothesis stating that difference in home victories and away team victories = 0 each year. We can set up a one-side t-test and reject the null if p-value is smaller than the significance level.

## 2. Has the total number of scored points per game changed over time?

```{r}
ggplot(data = avg_scores) +
  geom_line(aes(x = years, y = avg_totalscore), color = "#000000", size = 1.5) +
  geom_point(aes(x = years, y = avg_totalscore), color = "#000000", size = 3.0) +
  ggtitle("Average total scored points per year")

```

Other type of analysis for this conclusion: 
We can conduct a significance test for null hypothesis stating that scores each year is the same so expected value for each year's score = total scores of all years / number of years. We can set up a chi-square test and reject the null if test statistic is larger than corresponding critical score.
